### 2021-10-27 analysis/supplementary-materials/supplemental.Rmd

---
title: "Supplementary Materials"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  fig.height = 6,
  fig.width = 6,
  dpi = 300
)
library(cetaceanbcg)
library(tidyverse)
```

```{r motionless}
bw_elg <- bw180905_53_elg %>% 
  mutate(
    depth_min = map2_dbl(start, stop, function(t1, t2) {
      bw180905_53_10hz[between(bw180905_53_10hz$dt, t1, t2), "depth"] %>% min()
    }),
    depth_max = map2_dbl(start, stop, function(t1, t2) {
      bw180905_53_10hz[between(bw180905_53_10hz$dt, t1, t2), "depth"] %>% max()
    }),
    ygyr_min = map2_dbl(start, stop, function(t1, t2) {
      bw180905_53_10hz[between(bw180905_53_10hz$dt, t1, t2), ]$gw[,2] %>% min()
    }),
    ygyr_max = map2_dbl(start, stop, function(t1, t2) {
      bw180905_53_10hz[between(bw180905_53_10hz$dt, t1, t2), ]$gw[,2] %>% max()
    })
  )
bw10hz <- bw180905_53_10hz %>%
  mutate(
    rid_left = approx(bw_elg$start, 
                      bw_elg$region_id, 
                      dt, 
                      "constant")$y,
    rid_right = approx(bw_elg$stop, 
                       bw_elg$region_id, 
                       dt, 
                       "constant", 
                       yleft = 0)$y + 1,
    motion_class = case_when(
      depth < 2 ~ "surface",
      rid_left == rid_right ~ "motionless",
      TRUE ~ "active"),
    dive_id = split_dives(dt, depth,
                          surface = 2, min_depth = 10, min_dur = 5 * 60)
  )
```

```{r gyr_entropy, fig.cap="Motionless periods were manually identified by looking for periods of low amplitude rotational veloctiy ($\\omega$) around the lateral axis, indicative of cetaceans' dorso-ventral fluking motion."}
divesample <- filter(bw10hz, dive_id == 1)
diveelg <- bw_elg %>% 
  filter(start >= first(divesample$dt), stop <= last(divesample$dt))

depth_profile <- ggplot(divesample, aes(dt, depth)) +
  geom_line() +
  geom_rect(aes(xmin = start, xmax = stop, ymin = depth_min, ymax = depth_max),
            diveelg,
            inherit.aes = FALSE,
            fill = "orchid1",
            color = NA,
            alpha = 0.5) +
  scale_x_datetime(date_labels = "%H:%M:%S") +
  scale_y_reverse("Depth (m)") +
  theme_classic() +
  theme(axis.title.x = element_blank())
ygyr_profile <- ggplot(divesample, aes(dt, gw[, 2])) +
  geom_line() +
  geom_rect(aes(xmin = start, xmax = stop, ymin = ygyr_min, ymax = ygyr_max),
            diveelg,
            inherit.aes = FALSE,
            fill = "orchid1",
            color = NA,
            alpha = 0.5) +
  scale_x_datetime(date_labels = "%H:%M:%S") +
  scale_y_continuous(bquote(omega~(rad~s^{-1}))) +
  theme_classic() +
  theme(axis.title.x = element_blank())
cowplot::plot_grid(depth_profile, ygyr_profile, ncol = 1, align = "v")
```